---
title: "PLAN 372 - HW2"
author: "Anooshka Deshpande"
format: html
editor: visual
---

```{r}

library(tidyverse)
library(lubridate)
library(dbplyr)
```

```{r}
restaurant_data<-read_csv(here::here("restaurant_inspections.csv"))
view(restaurant_data)
```

Question 1:

```{r}
#I created a histogram because I want to observe the distribution of one quantitative variable, which is the inspection score.
ggplot(restaurant_data, aes(x = SCORE)) +
  geom_histogram()

```

Question 2:

```{r}
restaurant_age = mutate(restaurant_data, age_since_2024 = 2024 - year(restaurant_data$RESTAURANTOPENDATE))
ggplot(restaurant_age, aes(x = age_since_2024, y = restaurant_data$SCORE))+
  geom_point()
```

From this scatterplot, it looks like there is a weak positive relationship between restaurant age and score. I see that newer restaurants, particularly those under 10 years old, a higher amount of low inspection scores compared to older restaurants.

Question 3:

```{r}
sort(unique(restaurant_data$city_uppercase))
restaurant_data$CITY = str_to_upper(restaurant_data$CITY)
restaurant_data |>
  mutate(CITY = recode(CITY,
       "FUQUAY-VARINA" = "FUQUAY VARINA",
       "HOLLY SPRING" = "HOLLY SPRINGS",
       "MORRISVILE" = "MORRISVILLE",
       "RTP" = "RESEARCH TRIANGLE PARK"))

restaurant_data |>
  select(CITY, SCORE) |>
  group_by(CITY) |>
  summarise(mean = mean(SCORE))
```

`Yes, inspection scores vary by city. EVery city has a different mean inspection score.`

Question 4:

```{r}
#Question 4
#I used a boxplot because I want to observe whether a quantitative variable, score, varies with a categorical variable, inspector.
ggplot(restaurant_data, aes(x = INSPECTOR, y = SCORE))+
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Inspection scores vary by inspector, because the mean values and interquartile range on each of the individual boxplots are different. It is possible that a wider interquartile range corresponds to a more thorough inspector.

Question 5:

```{r}
restaurant_data |>
  group_by(CITY) |>
  summarise(n = n())
restaurant_data |>
  group_by(INSPECTOR) |>
  summarise(n = n())
```

The extreme results from the previous questions could be due to differences in sample size. I see that some towns, such as RTP and Wake Forest, underwent under 200 inspections, whereas other towns such as Raleigh underwent over 1800 inspections. I see that some inspectors conducted as few as 2 inspections, whereas others conducted nearly 350 inspections. This could introduce bias into the inspection scores.

Question 6:

```{r}
restaurant_data |>
  group_by(FACILITYTYPE) |>
  summarise(median = median(SCORE)) |>
  arrange(desc(median))
```

According to the table, the median score for restaurants is near the bottom of the range of median scores of all facility types. I can conclude that restaurants are not as clean as the other facility types.

Question 7:

```{r}
restaurant_only = restaurant_data |>
  filter(FACILITYTYPE == "Restaurant")
view(restaurant_only)
```
Step 1:
```{r}
ggplot(restaurant_only, aes(x = SCORE)) +
  geom_histogram()
```

Step 2:
```{r}
restaurant_only_age = mutate(restaurant_only, age_since_2024 = 2024 - year(restaurant_only$RESTAURANTOPENDATE))
ggplot(restaurant_only_age, aes(x = age_since_2024, y = restaurant_only$SCORE))+
  geom_point()
```
There is a weak positive correlation between age and inspection score.

Step 3:
```{r}
restaurant_only$CITY = str_to_upper(restaurant_only$CITY)
restaurant_only |>
  mutate(CITY = recode(CITY,
       "FUQUAY-VARINA" = "FUQUAY VARINA",
       "HOLLY SPRING" = "HOLLY SPRINGS",
       "MORRISVILE" = "MORRISVILLE",
       "RTP" = "RESEARCH TRIANGLE PARK"))

restaurant_only |>
  select(CITY, SCORE) |>
  group_by(CITY) |>
  summarise(mean = mean(SCORE))
```
Yes, inspection scores vary by city.

Step 4:
```{r}
ggplot(restaurant_only, aes(x = INSPECTOR, y = SCORE))+
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Yes, inspection scores vary by instructor. The mean values and interquartile range on each of the individual boxplots are different. It is possible that a wider interquartile range corresponds to a more thorough inspector.

Step 5:
```{r}
restaurant_only |>
  group_by(CITY) |>
  summarise(n = n())
restaurant_only |>
  group_by(INSPECTOR) |>
  summarise(n = n())
```
The sample size for each city and inspector vary by a large amount, and this could be an explanation for the extreme results seen in the previous questions.
